\chapter{Overview and Installation of DFTBopt}

\label{chap:install}

\section{Overview}

\textbf{DFTBopt}\cite{Vuong2018} is a package to automatically optimize
electronic, repulsive, and dispersion parameters for Density-Functional
based Tight-Binding (DFTB) method. The package includes two main programs:
(1) \textbf{repopt} for optimization of only repulsive potentials and (2)
\textbf{erepopt} for optimization of all DFTB parameters. \textbf{erepopt}
uses \textbf{repopt} for the repulsive potentials fitting. Currently,
\textbf{repopt} is stable released version 1.0. On the other hand,
\textbf{erepopt} is still in the beta-development version. In addition, the
package also provide some tools to analyze or evaluate new DFTB parameters.

Most of the codes were written in C++. The code required two libraries:
\href{http://lancet.mit.edu/ga/}{\textbf{galib247}} and
\href{http://eigen.tuxfamily.org/index.php?title=Main_Page}{\textbf{eigen3}}.
A part of the repulsive fitting (\textbf{repopt}) uses some piece of code
from a semi-automatic erepfit program originally developed by Michael
Gaus.\cite{Gaus2009} This manual was prepared using same style of DFTB+ manual,
making it looks similar to DFTB+ manual. $\ddot\smile\ddot\smile\ddot\smile$

\section{Installation}

To get the code:\\
{\scriptsize
git clone
\href{https://github.com/v2quan89/dftbopt.git}{https://github.com/v2quan89/dftbopt.git}
}

or download compressed file\\
{\scriptsize
wget \href{https://github.com/v2quan89/DFTBopt.tar.gz}{https://github.com/v2quan89/DFTBopt.tar.gz}
}

Note, both links may not work at this moment. You can send an email to
\textbf{v2quan89@gmail.com} if the links do not work.  Then, extract
DFTBopt.tar.gz using

{\scriptsize
tar -xzvf DFTBopt.tar.gz
}

Change to directory \textbf{DFTBopt} and type ``\textbf{./install.sh}''.
Currently, \textbf{DFTBopt} was tested for two compilers: GNU(g++) and
INTEL(icpc).

You can choose the compiler by setting \textbf{\$CXX=``g++''} or
\textbf{\$CXX=``icpc''} in the \textbf{``install.sh''} file.  To compile
\textbf{``erepopt''}, a MPI library is also required. The
\textbf{install.sh} script will try to get all the required library and
compile the code. You might have to adjust some flags in the
\textbf{makefile}. 

After the compilation, if you use \textbf{bash}, you can add following
command to you ``.bashrc'' or run it before using DFTBopt,

{\scriptsize
source path-to-DFTBopt-directory/dftbopt\_on.bash
}



